On exponential sums arising from the doubling method

### 2024-12-12 paper.tex
Stuck

\title{Degenerate Gauss sums for $\GL_n\left(\finiteField_q\right)$}

@@

Replacing $g$ with $h^{-1} g$, we arrive at the expression
\begin{align*}
	\dblGaussSum{\tau}{\chi} =& q^{-\frac{3n^2}{2}} \sum_{A \in \squareMatrix_n\left(\finiteField\right)} \fieldCharacter\left(\trace A\right) \sum_{h \in \GL_n\left(\finiteField\right)} \chi^{-1}\left(\det h\right) \tau\left(h^{-1}\right) \fieldCharacter\left(\trace\left(Ah\right)\right)\\
	& \circ \sum_{g \in \GL_n\left(\finiteField\right)} \tau\left(g\right) \fieldCharacter\left(-\trace Ag\right).
\end{align*}
Since the last inner sum is $q^{\frac{n^2}{2}} \GaussSum{\tau}{\fieldCharacter}_{-A}$, we have that the sum can be reduced to $A \in \GL_n\left(\finiteField\right)$. Replacing $h$ with $A^{-1} h$ and $g$ with $-A^{-1} g$, we get the sum
\begin{align*}
	\dblGaussSumScalar{\tau}{\chi} &= q^{-\frac{n^2}{2}} \centralCharacter{\tau}\left(-1\right) \sum_{A \in \GL_n\left(\finiteField\right)} \chi\left(\det A\right) \fieldCharacter\left(\trace A\right) \GaussSumScalar{\Contragradient{\tau} \times \chi^{-1}}{\fieldCharacter} \GaussSumScalar{\tau}{\fieldCharacter},
\end{align*}
which equals
\begin{align*}
	\dblGaussSumScalar{\tau}{\chi} &=  \centralCharacter{\tau}\left(-1\right) \GaussSumScalar{\chi}{\fieldCharacter}^n \GaussSumScalar{\Contragradient{\tau} \times \chi^{-1}}{\fieldCharacter} \GaussSumScalar{\tau}{\fieldCharacter}.
\end{align*}
This is probably also true if $1$ is in the cuspidal support of $\tau$ (return to this later: probably can follow from a multiplicativity arguments). In the meantime, we will check this for $\tau = 1$ the trivial representation of $\GL_1\left(\finiteField\right)$, in which case we have
$$\dblGaussSum{1}{\chi} = q^{-\frac{1}{2}} \sum_{\substack{x \in \multiplicativegroup{\finiteField} \\
		x \ne 1}} \chi\left(x - 1\right) = q^{-\frac{1}{2}}\left(-1 + \sum_{x \in \multiplicativegroup{\finiteField}} \chi\left(x\right) \right) = \begin{dcases}
		-q^{-\frac{1}{2}} & \chi \ne 1\\
		q^{-\frac{1}{2}} \left(q - 2\right) & \chi = 1
	\end{dcases}.$$

>>>>

\title{On exponential sums arising from the doubling method}

@@

Replacing $g$ with $h^{-1} g$, we arrive at the expression
\begin{align*}
	\dblGaussSum{\tau}{\chi} =& q^{-\frac{3n^2}{2}} \sum_{A \in \squareMatrix_n\left(\finiteField\right)} \fieldCharacter\left(\trace A\right) \sum_{h \in \GL_n\left(\finiteField\right)} \chi^{-1}\left(\det h\right) \tau\left(h^{-1}\right) \fieldCharacter\left(\trace\left(Ah\right)\right)\\
	& \circ \sum_{g \in \GL_n\left(\finiteField\right)} \tau\left(g\right) \fieldCharacter\left(-\trace Ag\right).
\end{align*}
Since the last inner sum is $q^{\frac{n^2}{2}} \GaussSum{\tau}{\fieldCharacter}_{-A}$, we have that the sum can be reduced to $A \in \GL_n\left(\finiteField\right)$. Replacing $h$ with $A^{-1} h$ and $g$ with $-A^{-1} g$, we get the sum
\begin{align*}
	\dblGaussSumScalar{\tau}{\chi} &= q^{-\frac{n^2}{2}} \centralCharacter{\tau}\left(-1\right) \sum_{A \in \GL_n\left(\finiteField\right)} \chi\left(\det A\right) \fieldCharacter\left(\trace A\right) \GaussSumScalar{\Contragradient{\tau} \times \chi^{-1}}{\fieldCharacter} \GaussSumScalar{\tau}{\fieldCharacter},
\end{align*}
which equals
\begin{align*}
	\dblGaussSumScalar{\tau}{\chi} &=  \centralCharacter{\tau}\left(-1\right) \GaussSumScalar{\chi}{\fieldCharacter}^n \GaussSumScalar{\Contragradient{\tau} \times \chi^{-1}}{\fieldCharacter} \GaussSumScalar{\tau}{\fieldCharacter}.
\end{align*}
This is probably also true if $1$ is in the cuspidal support of $\tau$  and $\chi \ne 1$ (return to this later: probably can follow from a multiplicativity arguments). In the meantime, we will check this for $\tau = 1$ the trivial representation of $\GL_1\left(\finiteField\right)$, in which case we have
$$\dblGaussSum{1}{\chi} = q^{-\frac{1}{2}} \sum_{\substack{x \in \multiplicativegroup{\finiteField} \\
		x \ne 1}} \chi\left(x - 1\right) = q^{-\frac{1}{2}}\left(-1 + \sum_{x \in \multiplicativegroup{\finiteField}} \chi\left(x\right) \right) = \begin{dcases}
		-q^{-\frac{1}{2}} & \chi \ne 1\\
		q^{-\frac{1}{2}} \left(q - 2\right) & \chi = 1
	\end{dcases}.$$
